---
title: "Two-way fixed effects estimation - Part 2: Solutions"
author: "Hans H. Sievertsen (h.h.sievertsen@bristol.ac.uk)"
date: "21. juni 2022"
output:
  beamer_presentation: default
  slidy_presentation: default
theme: metropolis
latex_engine: xelatex
fontsize: 8pt
---



```{r , include=FALSE, echo=FALSE}
knitr::opts_chunk$set(dpi=300,echo = FALSE, cache = TRUE, warning=F, message = FALSE)
library("tidyverse")
library("ggpubr")
library("bacondecomp")
library("lubridate")
library("fixest")
library("patchwork")
library("DIDmultiplegt")
library("scales")
library("readstata13")
```

```{r echo=FALSE, cache=T}

#df%>%filter(G!=1)%>%mutate(true=y-1)%>%select(true)%>%colMeans()
simulate_data<-function(seed,T2,T3,m2,m3,mt2,mt3,G2,G3){
  # set values
  set.seed(seed)                        # set seed
  mt2=(1+mt2)                           # Treatment effect growth group 2
  mt3=(1+mt3)                           # Treatment effect growth group 3
  m2=1+m2                               # Treatment effect group 2
  m3=1+m3                               # Treatment effect group 3
  T<-50                                 # Time periods
  G<-3                                  # Groups
  G1<-25                                # Group size group 1 (never treated)
  N<-G1+G2+G3   
  # Create tibble
  df<-tibble(id=rep(1:(N),T),                                     # Id variable 1 2 3 ... 1 2 3 
             t=rep(1:T,each=(N)))%>%                              # Time variable 1 1 1 1 .... 2 2 2
             mutate(G=ifelse(id<=G1,1,ifelse(id>G1&id<=(G1+G2),2,3)),    # Treated: D==1
             D=ifelse(G==2&t>=T2,1,ifelse(G==3&t>=T3,1,0)),         # Post treatment indicator 
             mean=ifelse(D==1&G==2,m2*mt2^(t-T2),                 # Treatment effect group 2 
                  ifelse(D==1&G==3,m3*mt3^(t-T3),                 # Treatment effect group 3
                              1)),                                # mean Y for untreated
                          mean=ifelse(G==1,0.25,mean),
              y=rnorm(n=N*T,mean=mean,sd=.1),                  # Simulate Outcome y
           section=case_when(
             t<5~1,
             t>4&t<25~2,
             t>24~3),
           section1=case_when(
             t<5~1,
             t>4&t<30~2,
             t>29~3))%>%                  # Simulate Outcome y
    group_by(G,D)%>%                                              # Group by "group times post" treatment
    mutate(ybar=mean(y))%>%                                          # Group means for chart
    group_by(G,section)%>%                                              # Group by "group times post" treatment
    mutate(ybar1=mean(y))%>%                                         # Group means for chart
      group_by(G,section1)%>%                                              # Group by "group times post" treatment
    mutate(ybar2=mean(y))                                          # Group means for chart
  
  df<-df%>%mutate(Date=ymd("2009-12-31") %m+% months(t),
                kom=factor(case_when(
                  G==1~"Tønder",
                  G==2~"Sønderborg",
                  G==3~"Aabenraa"                ),
                  levels=c("Tønder","Aabenraa","Sønderborg")))
  
  return(df)
}
df<-simulate_data(1909,5,30,1,1,0.03,0.00,100,50)
analysisdata<-df%>%filter(G!=3,t<30)%>%mutate(after=ifelse(t>4,1,0),treated=ifelse(G==2,1,0),
                                              afterXtreated=after*treated,m=G)
analysisdata_update<-df%>%mutate(after=ifelse(t>4,1,0),treated=ifelse(G==2,1,0),
                                              afterXtreated=after*treated,m=G)
save.dta13(analysisdata_update%>%mutate(true=ifelse(D==1,mean-1,NA))%>%ungroup()%>%select(id,t,G,D,y,true),"analysisdata.dta")
dfcd<-df%>%group_by(kom,Date)%>%summarise(D=mean(D),y=mean(y),G=mean(G),t=mean(t))

```



# Plan for today

1. Recap: Two-way fixed effects $\neq$ difference-in-differences.

2. Apply solutions in Stata on simulated and real example. 


- I will share do files, slides, and zoom recording.


---


# Recap: Difference-in-Differences - Chart

- Aabenraa Kommune is treated with an intervention in June 2012 (*D=1*). 
- Tønder Kommune is never treated (*D=0*)
- Outcome of interest *y* 
- Data for individuals in Sønderborg (N=50 individuals) and Tønder (N=25 individuals) for the period January 2010 to December 2011.
 
 
```{r , echo = FALSE,fig.width=3.5,fig.height=2.5,fig.align="center"}
ggplot(df%>%filter(G!=2))+geom_jitter(aes(x=Date,y=y,colour=as.factor(kom)),size=.05,alpha=0.5)+theme_classic()+ theme(legend.position="top")+
  labs(x="t",y="y",colour="g") +scale_color_manual(values=c("#999999", "#56B4E9"))+ylim(0,4)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
  geom_vline(xintercept = 10, linetype="dashed", color = "gray", size=0.5)+
  labs(x=" ",color=" " )+theme(text=element_text(size=11))+
  guides(color = guide_legend(override.aes = list(size = 2)))

```




---


# Recap: Difference-in-Differences - Means


```{r, echo = FALSE,fig.width=3.5,fig.height=2.5,fig.align="center"}
ggplot()+
  geom_jitter(df%>%filter(G!=2),mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=.05,alpha=0.15)+theme_classic()+ theme(legend.position="top")+
  geom_line(df%>%filter(G!=2,t<30),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=1)+
  geom_line(df%>%filter(G!=2,t>=30),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=1)+
  labs(x="t",y="y",colour="g") +scale_color_manual(values=c("#999999",  "#56B4E9"))+ylim(0,4)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
  geom_vline(xintercept = 10, linetype="dashed", color = "gray", size=0.5)+
  labs(x=" ",color=" " )+theme(text=element_text(size=11))+
  guides(color = guide_legend(override.aes = list(size = 2)))+
  geom_text(aes(x=ymd("2013-03-31"),y=2.25,label="2.00"),color="#56B4E9",size=3)+
  geom_text(aes(x=ymd("2011-01-15"),y=1.25,label="1.00"),color="#56B4E9",size=3)+
  geom_text(aes(x=ymd("2013-03-31"),y=0.05,label=0.25),color="#999999",size=3)+
  geom_text(aes(x=ymd("2011-01-31"),y=0.05,label=0.25),color="#999999",size=3)

#summary(df%>%ungroup()%>%filter(t<30,G==3)%>%select(y))
#summary(df%>%ungroup()%>%filter(t>=30,G==3)%>%select(y))
#summary(df%>%ungroup()%>%filter(t<30,G==1)%>%select(y))
#summary(df%>%ungroup()%>%filter(t>=30,G==1)%>%select(y))
```

- Difference Aabenraa: 2.00-1.00=1.00
- Difference Tønder: 0.25-0.25=0.00
- Difference-in-Differences=1.00-0.00=1.00


---

# Recap: Difference-in-Differences - Regression

We use OLS to estimate

$y=\beta_0+\beta_1 treated +\beta_{2} after + \beta_{DiD} after \times treated+u$

- $after$ 1 if June 2012 or later, 0 otherwise.
- $treated$ 1 if Aabenraa, 0 otherwise.


```{r, echo=F, include=F}
analysisdata<-analysisdata_update%>%filter(G!=2)%>%
  mutate(after=ifelse(t>=30,1,0),treated=ifelse(G==3,1,0),afterXtreated=after*treated)
```

```{r, echo=T}
feols(y~treated+after+afterXtreated,data=analysisdata)

```





---

# Recap: Difference-in-Differences - TWFE


We use OLS to estimate the Two-Way Fixed Effects (TWFE) model

$y=\alpha+\beta_{TWFE} D+{\tau}'{t}+{\mu}'{m}+u$


- ${t}$ a vector of month dummies
- ${m}$ a vector of municipality dummies
- $D$ =1 if treated, 0 otherwise

```{r, echo=T}
feols(y~D|t+m,data=analysisdata)

```



---

# Recap: More data - More problems


- Sønderborg Kommune is treated with the same intervention in May 2010 and onwards. 

```{r , echo = FALSE,fig.width=3.5,fig.height=2.5,fig.align="center"}
ggplot(df)+geom_jitter(aes(x=Date,y=y,colour=as.factor(kom)),size=.05,alpha=0.15)+theme_classic()+ theme(legend.position="top")+
  labs(x="t",y="y",colour="g") +scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+ylim(0,8)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
  geom_vline(xintercept = 10, linetype="dashed", color = "gray", size=0.5)+
  labs(x=" ",color=" " )+theme(text=element_text(size=11))+
  guides(color = guide_legend(override.aes = list(size = 2)))

```


---

# Recap: TWFE with more groups

We again use OLS to estimate the Two-Way Fixed Effects (TWFE) model

$y=\alpha+\beta_{TWFE} D+{\tau}'{t}+{\mu}'{m}+u$


- ${t}$ a vector of month dummies
- ${m}$ a vector of municipality dummies
- $D$ =1 if in Sønderborg May 2010 or later, 1 if in Aabenraa in June 2012 or later, 0 otherwise


```{r, echo=T}
feols(y~D|t+m,data=analysisdata_update)

```



---

# Recap: How does $\beta_{TWFE}$ relate to $\beta_{DiD}$?

**What is ${\beta}_{TWFE}$ actually capturing?**
  \vspace{15pt}

- *Goodman-Bacon (GB): "Difference-in-Differences with Variation in Treatment Timing" (2021, JoE )*

  - **Decomposition of $\hat{\beta}_{TWFE}$  in weighted $\hat{\beta}_{DiD}$s**
  - Applicable to staggered adoption designs.
  
  
- *Chaisemartin & D'Haultfoeuille (CD): "Two-way fixed effects estimators with heterogeneous treatment effects" (2020, AER)*

  - **Decomposition of $\hat{\beta}_{TWFE}$  in weighted TEs across $(t,m)$ cells.**
  - Applicable to 2-way (e.g., group & time) fixed effects approaches.
  
  


\vspace{15pt}
  



---

# Recap: The Goodman-Bacon Decomposition

- Goodman Bacon: we can decompose 

```{r , echo = FALSE,fig.width=3,fig.height=2,fig.align="center"}
ggplot()+geom_jitter(df,mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=.05,alpha=0.1)+theme_classic()+ theme(legend.position="top")+
  labs(x="t",y="y",colour="g") +scale_color_manual(values=c("#999999","#56B4E9", "#E69F00"))+ylim(0,8)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
  geom_vline(xintercept = 10, linetype="dashed", color = "gray", size=0.5)+
  labs(x=" ",color=" " )+theme(text=element_text(size=9))+
  guides(color = guide_legend(override.aes = list(size = 2)))+
  geom_vline(xintercept=ymd("2010-05-15"),alpha=.1)+
  geom_vline(xintercept=ymd("2012-06-15"),alpha=.1)

```

- into four 2X2 DiDs:

```{r , echo = FALSE,fig.width=4,fig.height=2,fig.align="center"}

bp<-ggplot(df)+theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour="g")+
  scale_x_date(date_breaks = "25 months", 
                 labels=date_format("%b-%y"),
                 limits = as.Date(c('2010-01-01','2014-03-01')))+ylim(0,7)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=4),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))


a<-bp+geom_step(data=df%>%filter(G!=3),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999","#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))
b<-bp+geom_step(data=df%>%filter(G!=2),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999",  "#56B4E9"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))
c<-bp+geom_step(data=df%>%filter(G!=1,t<30),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))
d<-bp+geom_step(data=df%>%filter(G!=1,t>5),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c( "#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))

ggarrange(a,b,c,d,labels = c("DiD1: Sønd (T) vs Tønd (C) ", "DiD2: Aabe (T) vs Tønd (C)", "DiD3: Sønd (T) vs Aabe (C) ","DiD4: Aabe (T) vs Sønd (C)"),ncol = 4, nrow = 2,common.legend = TRUE, legend = "bottom", font.label = list(size = 3, face = "plain", color ="black"))
```





---

# Recap: DiD4 the Bad Guy!


```{r , echo = FALSE,fig.width=3,fig.height=2, fig.align="center"}

bp<-ggplot(df)+theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour="g")+
  scale_x_date(date_breaks = "12 months", 
                 labels=date_format("%b-%y"),
                 limits = as.Date(c('2010-01-01','2014-03-01')))+ylim(0,7)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=4),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))


a<-bp+geom_step(data=df%>%filter(G!=3),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.5)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999", "#56B4E9","#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=14))
b<-bp+geom_step(data=df%>%filter(G!=2),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.5)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999",  "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=14))
c<-bp+geom_step(data=df%>%filter(G!=1,t<30),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.5)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=14))
d<-bp+geom_jitter(data=df%>%filter(G!=1,t>5),mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=0.1,alpha=0.1)+geom_step(data=df%>%filter(G!=1,t>5),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=1)+
  geom_vline(xintercept = ymd("2012-06-15"), linetype="dashed", color = "gray", size=1.25)+
  scale_color_manual(values=c( "#56B4E9", "#E69F00"))+
  labs(x=" ",color=" ",title="DiD4: Aabe (T) vs Sønd (C)")+theme(text=element_text(size=7))


d

#summary(analysisdata_update%>%ungroup()%>%filter(t<30,G==3,t>4)%>%select(y))
#summary(analysisdata_update%>%ungroup()%>%filter(t>=30,G==3)%>%select(y))
#summary(analysisdata_update%>%ungroup()%>%filter(t<30,G==2)%>%select(y))
#summary(analysisdata_update%>%ungroup()%>%filter(t>=30,G==2)%>%select(y))

```


- Difference Aabenraa (Treated): 2.00-1.00=1.00
- Difference Sønderborg (Always Treated: Our Control!):  5.72-2.65=3.07
- Difference-in-Differences: 1.00-3.07=-2.00

We use Sønderborg as a control group, because it doesn't change treatment status. However, because of dynamic treatment effects (for Sønderborg), Sønderborg is a poor control because the number of periods it has been treated changes over time! 

- Conclusion: Trend for Sønderborg is not a good counterfactual for Aabenraa!




---


# Recap: The Goodman-Bacon ${\beta}_{TWFE}$ decomposition

${\beta}_{TWFE}$ ***is the weighted average across these 4 DiDs***:

$$\hat{\beta}_{TWFE}=w_1\hat{\beta}_{DiD1}+w_2\hat{\beta}_{DiD2}+w_3\hat{\beta}_{DiD3}+w_4\hat{\beta}_{DiD4}$$


- Key insight: $w\neq$ population shares, but also **depends on when a group gets treated**. 

- See Theorem 1 in GB for the general expression of (3) (equation 10a in his paper) and the definition of the weights. 



# Recap: Goodman-Bacon Weights in our Example

  
```{r , echo = FALSE,include=FALSE,cache=T}
df_bacon <- bacon(y ~ D,data = df,id_var = "id",time_var = "t")    # Bacon decomp
cc<-feols(y ~ D | G+ t, df)
```



```{r , echo = FALSE,fig.width=4,fig.height=1,fig.align="center"}


bp<-ggplot(df)+theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour="g")+
  scale_x_date(date_breaks = "25 months", 
                 labels=date_format("%b-%y"),
                 limits = as.Date(c('2010-01-01','2014-03-01')))+ylim(0,7)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=4),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))


a<-bp+geom_step(data=df%>%filter(G!=3),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999",  "#56B4E9"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))
b<-bp+geom_step(data=df%>%filter(G!=2),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999",  "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))
c<-bp+geom_step(data=df%>%filter(G!=1,t<30),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))
d<-bp+geom_step(data=df%>%filter(G!=1,t>5),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c( "#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=6))

ggarrange(a,b,c,d,labels = c(paste("Sø (T) vs Tø (C): w1=",round(df_bacon$weight[1],digits=2)),paste("Ab (T) vs Tø (C): w2=",round(df_bacon$weight[2],digits=2)),paste("Sø (T) vs Ab (C): w3=",round(df_bacon$weight[4],digits=2)),paste("Ab (T) vs Sø (C): w4=",round(df_bacon$weight[3],digits=2))),ncol = 4,  legend = "bottom", font.label = list(size = 4, face = "plain", color ="black"),nrow = 1,common.legend = TRUE )

```

**Recap: Decomposing the TWFE** 

$\hat{\beta}_{TWFE}=$ `r round(df_bacon$weight[1],digits=3)` $\times$ `r round(df_bacon$estimate[1],digits=3)`+ 
  `r round(df_bacon$weight[2],digits=3)` $\times$ `r round(df_bacon$estimate[2],digits=3)`+ 
  `r round(df_bacon$weight[4],digits=3)` $\times$ `r round(df_bacon$estimate[4],digits=3)`+ 
  `r round(df_bacon$weight[3],digits=3)` $\times$ `r round(df_bacon$estimate[3],digits=3)`=`r round(as.numeric(cc$coefficients[1]),digits=2)`

- Weight Sønderborg:0.11+0.12=0.23. $N_{\text{Sø}}=100$
- Weight Aabenraa:0.18+0.60=0.78. $N_{\text{Ab}}=50$
- Aabenraa gets a larger weight because it is treated more in the middle.
- Not necessarily a problem if you know that it is not population weighted!





---

# Recap: The CD decomposition

**Decompose $\beta_{TWFE}$  in weighted TEs across  $(m,t):D_{m,t}=1$ cells**

- A ${m,t}$ cell is time (one of the fixed effects) period times area (the other fixed effect) treated unit. 
   
- $\beta_{TWFE}$ is then given by the weighted average across all these cells.

$$\beta_{TWFE}=w_{Sønderborg,May2010} TE_{S,5:2010}+w_{S,6:2010} TE_{S,6:2010}...$$ +

- (This is more "general" than GB)

- Cells that are treated when no one else is treated => large weight
- A treated cell for a group that is rarely treated => large weight!




---

# Recap: The CD   decomposition

**Decompose $\beta_{TWFE}$  in weighted TEs across  $(m,t):D_{m,t}=1$ cells**


```{r , echo = FALSE,fig.width=3.5,fig.height=2.5,fig.align="center"}


ggplot()+
  geom_point(dfcd,mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=1)+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +scale_color_manual(values=c("#999999","#56B4E9", "#E69F00"))+ylim(0,8)+
   theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
   geom_vline(xintercept = ymd("2010-05-15"), linetype="dashed", color = "gray", size=0.5)+
  geom_vline(xintercept = ymd("2012-06-15"), linetype="dashed", color = "gray", size=0.5)+theme(text=element_text(size=11))+
    guides(color = guide_legend(override.aes = list(size = 2)))+
  geom_point(data=dfcd%>%filter(t==6&G==2),mapping=aes(x=Date,y=y),color="black",size=2,shape=1)+
  geom_text(aes(x=ymd("2010-08-31"),y=1.5),label="(m=Sø,t=June 2010)", size=2)

```




---

# Recap: The CD   decomposition

**Decompose $\beta_{TWFE}$  in weighted TEs across  $(m,t):D_{m,t}=1$ cells**


```{r , echo = FALSE,fig.width=3.5,fig.height=2.5,fig.align="center"}

ggplot()+
  geom_point(dfcd,mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=1)+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +scale_color_manual(values=c("#999999", "#56B4E9","#E69F00"))+ylim(0,8)+
   theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
   geom_vline(xintercept = ymd("2010-05-15"), linetype="dashed", color = "gray", size=0.5)+
  geom_vline(xintercept = ymd("2012-06-15"), linetype="dashed", color = "gray", size=0.5)+theme(text=element_text(size=11))+
    guides(color = guide_legend(override.aes = list(size = 2)))+
  geom_point(data=dfcd%>%filter(t==50&G==2),mapping=aes(x=Date,y=y),color="black",size=2,shape=1)+
  geom_text(aes(x=ymd("2013-07-31"),y=7.8),label="(m=Sø,t=Feb 2014)", size=2)

```

\addtocounter{framenumber}{-1}
  
  


  
  
---

# Recap: The CD weights in our example

```{r , echo = FALSE,fig.width=3,fig.height=2,fig.align="center",warning=FALSE}
m<-lm(D~as.factor(t)+as.factor(G),data=df)
df$res<-m$residuals
# calculate weights
w<-df%>%filter(G!=1)%>%group_by(Date,G)%>%summarise(d=mean(D),res=mean(res),y=mean(y),n=sum(D))%>%filter(d==1)%>%
  ungroup()%>%mutate(weight=(n/sum(n))*res/(sum((n/sum(n))*res)))

# charts
ggplot(dfcd)+geom_point(aes(x=Date,y=y,colour=kom),show.legend = F,size=1)+
  scale_color_manual(values=c("#999999", "#56B4E9","#E69F00"))+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +ylim(0,8)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
  geom_point(data=w,mapping=aes(x=Date,y=y+0.2,shape=as.factor(round(weight,digits=6))),size=1)+labs(colour="g",shape="weight")+theme(text=element_text(size=9))+
  guides(color = guide_legend(override.aes = list(size = 2)))


```

- Sønderborg cells  get  negative weight after $t=30$ because they are from a group that is mostly treated and at a time that that is mostly treated.
- Note that in contrast to GB, we cannot empirically decompose $\beta_{TWFE}$ because we don't know the TEs in the $m,t$ cells!



  

  
---

# Recap: Linking GB and CD


**GB weights**

```{r , echo = FALSE,fig.width=4,fig.height=1,fig.align="center"}


bp<-ggplot(df)+theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour="g")+
  scale_x_date(date_breaks = "25 months", 
                 labels=date_format("%b-%y"),
                 limits = as.Date(c('2010-01-01','2014-03-01')))+ylim(0,7)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=4),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))


a<-bp+geom_step(data=df%>%filter(G!=3),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999", "#56B4E9"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=5))
b<-bp+geom_step(data=df%>%filter(G!=2),mapping=aes(x=Date,y=ybar,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#999999",  "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=5))
c<-bp+geom_step(data=df%>%filter(G!=1,t<30),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2010-05-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=5))
d<-bp+geom_step(data=df%>%filter(G!=1,t>5),mapping=aes(x=Date,y=ybar2,colour=as.factor(kom)),size=.75)+
  geom_vline(xintercept = ymd("2012-06-01"), linetype="dashed", color = "gray", size=0.25)+
  scale_color_manual(values=c( "#56B4E9", "#E69F00"))+
  labs(x=" ",color=" " )+theme(text=element_text(size=5))

ggarrange(a,b,c,d,labels = c(paste("Sø (T) vs Tø (C): w1=",round(df_bacon$weight[1],digits=2)),paste("Ab (T) vs Tø (C): w2=",round(df_bacon$weight[2],digits=2)),paste("Sø (T) vs Ab (C): w3=",round(df_bacon$weight[4],digits=2)),paste("Ab (T) vs Sø (C): w4=",round(df_bacon$weight[3],digits=2))),ncol = 4, nrow = 1,common.legend = T, legend = "bottom", font.label = list(size = 3, face = "plain", color ="black"))

```
\vspace{-18pt}
- Weight Sønderborg:0.11+0.12=0.23.
- Weight Aabenraa:0.18+0.60+0.78. 

**CD weights**



```{r , echo = FALSE,fig.width=2.5,fig.height=1.5,fig.align="center",warning=FALSE}


# charts
ggplot(dfcd)+geom_point(aes(x=Date,y=y,colour=kom),show.legend = F,size=.5)+
  scale_color_manual(values=c("#999999",  "#56B4E9","#E69F00"))+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +ylim(0,8)+
  theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        text = element_text(size=5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA"),legend.margin=margin(b=-10),legend.box.margin=margin(0,0,0,0))+
  geom_point(data=w,mapping=aes(x=Date,y=y+0.2,shape=as.factor(round(weight,digits=6))),size=.5)+labs(colour="g",shape="weight")+theme(text=element_text(size=8))+
  guides(color = guide_legend(override.aes = list(size = 1)))

```

\vspace{-18pt}
- Weight Sønderborg:$-0.026\times 21+0.031\times 25$ =0.23.
- Weight Aabenraa:$0.037\times 21$=0.78.



---

# What to do?

**So you have a TWFE analysis. Ask: **

1. Are treatment effects homogeneous?
2. Are treatment effects dynamic?
3. Do you have negative weights?


# Solutions: A Roadmap:

1. Ruling out dynamic and heterogeneous treatments effects a priori is often a strong assumption. So I typically assume that could be the case. 

2. Calculate weights (GB and/or CD)

3. Use estimator that handles the weight issue and (potentially) allows for dynamic effects.




---

# Solution 1: CD's  estimator

\begin{columns}
\begin{column}{0.5\linewidth}

\begin{itemize}
\item  Chaisemartin \& D'Haultfoeuille (CD): (2020, AER) 
\item  Not allowing dynamic effects!
\item  \alert{Compares adjacent periods for switching in and out cells.}
\item  Switching in (joiners) relies on // assumption on untreated outcomes.
\item Switching out (leavers) relies on // assumption on treated outcomes.

\end{itemize}
\end{column}





\begin{column}{0.5\linewidth}

```{r , echo = FALSE,fig.width=2,fig.height=2,fig.align="center"}

ggplot()+
  geom_point(dfcd,mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=.5)+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +scale_color_manual(values=c("#999999",  "#56B4E9","#E69F00"))+ylim(0,8)+
   theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         text = element_text(size=1.5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         panel.background = element_rect(fill = "#FAFAFA"), legend.position = "none")+
   geom_vline(xintercept = ymd("2010-05-15"), linetype="dashed", color = "gray", size=0.15)+
  geom_vline(xintercept = ymd("2012-06-15"), linetype="dashed", color = "gray", size=0.5)+theme(text=element_text(size=11))+
    guides(color = guide_legend(override.aes = list(size = 1.75)))

```

\end{column}
\end{columns}


---

# Allowing for dynamic effects

- Goal: Estimate event-study design with OLS for effects at $l$ periods to treatment.

$$y_{g,t}=\alpha_{g}+\tau_{t}+\sum \beta_{TWFE,l}1\{l==1\}+e_{g,t}
$$


- This approach also suffers from the issues listed above,

$$E[\hat{\beta}_{TWFE,l}]=\sum w_{g,l}TE_{g,l}+\sum\sum w_{g,l'}TE_{g,l'}$$
  
1. First sum might have negative weights (very similar as before)!
2. And also contamination from other periods $l'\neq l$ treatment effect.

- Decide on how to aggregate effects across groups!

(see Sun and Abraham (2021) and Chaisemartin's "Advances in Difference-in-Differences in Econometrics" talk in December 2022).
  
---

# Solution 2: Callaway and Sant'Anna (2021, JE) estimator



\begin{columns}
\begin{column}{0.5\linewidth}

\begin{itemize}
\item [] (Sun \& Abraham (2021, JE) is very much of the same spirit)
\item  Create groups, $g$, that start treatment at the same time, $c$.
\item  To get the effect of having been treated for $l$ periods:
\item []
\begin{itemize}
\item []
\item[1.] Compute difference in $y$ for group $c$ between period $c+l$ and period $c-1$.
\item[2.]  Compute average difference in $y$ for never treated groups between period $c+l$ and period $c-l$. 
\item[3.]  Compute difference between 1 and 2.
\end{itemize}
\item []
\item Key assumption: // based on never having been treated!
\end{itemize}
\end{column}


\begin{column}{0.5\linewidth}

```{r , echo = FALSE,fig.width=2,fig.height=2,fig.align="center"}

ggplot()+
  geom_point(dfcd,mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=.5)+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +scale_color_manual(values=c("#999999",  "#56B4E9","#E69F00"))+ylim(0,8)+
   theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         text = element_text(size=1.5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         panel.background = element_rect(fill = "#FAFAFA"), legend.position = "none")+
   geom_vline(xintercept = ymd("2010-05-15"), linetype="dashed", color = "gray", size=0.15)+
  geom_vline(xintercept = ymd("2012-06-15"), linetype="dashed", color = "gray", size=0.5)+theme(text=element_text(size=11))+
    guides(color = guide_legend(override.aes = list(size = 1.75)))

```

\end{column}
\end{columns}



---

# Solution 3: Borusyak, Jaravel, and Spiess (BJS) (2022)

\begin{columns}
\begin{column}{0.5\linewidth}

\begin{itemize}
\item As CS, but to get the effect of having been treated for $l$ periods:
\item []
\begin{itemize}
\item []
\item[1.] Compute difference in $y$ for group $c$ between period $c+l$ and average across all periods  $t=0$ and  $c-1$.
\item [2.] Compute average difference in $y$ for never treated groups between period $c+l$ and  average across all periods  $t=0$ and  $c-1$. 
\item [3.] Compute difference between 1 and 2. 
\end{itemize}
\item []
\item More efficient than CS because it uses more data.
\item But more sensitive to violation of common trends assumption!
\item Bias-variance trade-off
\end{itemize}
\end{column}


\begin{column}{0.5\linewidth}

```{r , echo = FALSE,fig.width=2,fig.height=2,fig.align="center"}

ggplot()+
  geom_point(dfcd,mapping=aes(x=Date,y=y,colour=as.factor(kom)),size=.5)+
  theme_classic()+ theme(legend.position="top")+
  labs(x=" ",y="y",colour=" ") +scale_color_manual(values=c("#999999",  "#56B4E9","#E69F00"))+ylim(0,8)+
   theme(axis.ticks = element_blank(),plot.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         text = element_text(size=1.5),legend.background = element_rect(fill="#FAFAFA",colour = "#FAFAFA"),
         panel.background = element_rect(fill = "#FAFAFA"), legend.position = "none")+
   geom_vline(xintercept = ymd("2010-05-15"), linetype="dashed", color = "gray", size=0.15)+
  geom_vline(xintercept = ymd("2012-06-15"), linetype="dashed", color = "gray", size=0.5)+theme(text=element_text(size=11))+
    guides(color = guide_legend(override.aes = list(size = 1.75)))

```

\end{column}
\end{columns}

---

Let's try them in Stata on our simulated exampel and on a real example

---

# A real example


\begin{columns}
\begin{column}{0.5\linewidth}
\includegraphics[width=0.9\linewidth]{fig4.png}
\end{column}
\begin{column}{0.5\linewidth}
Valente, Sievertsen, \& Puri (2021)
\begin{itemize}
\item Staggered rollout of chlorhexidine gel (CHX) treatment across districts in Nepal. 
\item Estimate effect of CHX on mortality using district and month of fixed effects.
\end{itemize}
\end{column}
\end{columns}


